The unequal distribution of access to education and transport: insights from for São Paulo

### 2016-02-11 paper.tex

\documentclass[]{article}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \ifxetex
    \usepackage{mathspec}
    \usepackage{xltxtra,xunicode}
  \else
    \usepackage{fontspec}
  \fi
  \defaultfontfeatures{Mapping=tex-text,Scale=MatchLowercase}
  \newcommand{\euro}{€}
\fi
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
% use microtype if available
\IfFileExists{microtype.sty}{%
\usepackage{microtype}
\UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\usepackage[margin=1in]{geometry}
\ifxetex
  \usepackage[setpagesize=false, % page size defined by xetex
              unicode=false, % unicode breaks when used with xetex
              xetex]{hyperref}
\else
  \usepackage[unicode=true]{hyperref}
\fi
\hypersetup{breaklinks=true,
            bookmarks=true,
            pdfauthor={Ana I. Moreno-Monroy (Universitat Rovira i Virgili), Frederico R. Ramos (Fundação Getulio Vargas) and Robin Lovelace (University of Leeds)},
            pdftitle={The unequal distribution of access to education and transport: insights from for São Paulo},
            colorlinks=true,
            citecolor=blue,
            urlcolor=blue,
            linkcolor=magenta,
            pdfborder={0 0 0}}
\urlstyle{same}  % don't use monospace font for urls
\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 2pt minus 1pt}
\setlength{\emergencystretch}{3em}  % prevent overfull lines
\setcounter{secnumdepth}{0}

@@

\begin{document}

\maketitle

@@

\begin{itemize}
\itemsep1pt\parskip0pt\parsep0pt
\item
  Motivations
\item
  General measures of accessibility
\item
  School accessibility
\end{itemize}

\ldots{}

@@

@@

The cumulative-opportunity measure for mode \(M\) (Boisjoly and
El-Geneidy, 2016) is defined as:

@@

\begin{itemize}
\itemsep1pt\parskip0pt\parsep0pt
\item
  Descriptive statistics about modal choice and travel time by AEPs and
  level of income
\item
  Spatial Distribution of Schools and Transit-Dependent Students at the
  AEP Level
\end{itemize}

@@

Allen, W Bruce, Dong Liu, and Scott Singer. 1993. ``Accesibility
Measures of US Metropolitan Areas.'' \emph{Transportation Research Part
B: Methodological} 27 (6). Elsevier: 439--49.

Ingram, David R. 1971. ``The Concept of Accessibility: A Search for an
Operational Form.'' \emph{Regional Studies} 5 (2). Taylor \& Francis:
101--7.

>>>>

\documentclass[]{article}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \ifxetex
    \usepackage{mathspec}
    \usepackage{xltxtra,xunicode}
  \else
    \usepackage{fontspec}
  \fi
  \defaultfontfeatures{Mapping=tex-text,Scale=MatchLowercase}
  \newcommand{\euro}{€}
\fi
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
% use microtype if available
\IfFileExists{microtype.sty}{%
\usepackage{microtype}
\UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\usepackage[margin=1in]{geometry}
\ifxetex
  \usepackage[setpagesize=false, % page size defined by xetex
              unicode=false, % unicode breaks when used with xetex
              xetex]{hyperref}
\else
  \usepackage[unicode=true]{hyperref}
\fi
\hypersetup{breaklinks=true,
            bookmarks=true,
            pdfauthor={Ana I. Moreno-Monroy (Universitat Rovira i Virgili), Frederico R. Ramos (Fundação Getulio Vargas) and Robin Lovelace (University of Leeds)},
            pdftitle={The unequal distribution of access to education and transport: insights from for São Paulo},
            colorlinks=true,
            citecolor=blue,
            urlcolor=blue,
            linkcolor=magenta,
            pdfborder={0 0 0}}
\urlstyle{same}  % don't use monospace font for urls
\usepackage{graphicx,grffile}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 2pt minus 1pt}
\setlength{\emergencystretch}{3em}  % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{0}

@@

% Redefines (sub)paragraphs to behave more like sections
\ifx\paragraph\undefined\else
\let\oldparagraph\paragraph
\renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
\let\oldsubparagraph\subparagraph
\renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi

\begin{document}
\maketitle

@@

\begin{itemize}
\tightlist
\item
  Motivations
\item
  General measures of accessibility
\item
  School accessibility
\item
  Literature review
\end{itemize}

References transport and social exclusion (Hernandez and Titheridge
2015) and (Hernandez and Titheridge 2015) provides a general framework
to undertand transport-related social exclusion in emerging economies

Evidence on access inequalities in the Brazilian context (Maia et al.
2016)

Active travel children and adolescents in the SPMR (Sá et al. 2015)
\ldots{}

@@

\section{Area of study}\label{area-of-study}

The São Paulo Metropolitan Region (SPMR) is a large metropolis located
in the South West of Brazil. It extends for over 7,700 squared
kilometers and groups 39 different minucipalities. The estimated
population in 2010 was close to 21 million. \ldots{} ADD MAP, etc.

\section{Data}\label{data}

For building our school accessibility index, we use data from three
different sources. The first one is the 2008 School Census of the
Brazilian National Institute of Educational Studies and Surveys (INEP),
which provides information on the (universe) of public educational
institutions, including their exact location (postal address) and the
number of students enrolled in primary and secondary education. A
clarification about the educational system in Brazil is in place here.
Primary education \textit{ensino fundamental} comprises nine years,
divided on a first year of basic literacy
\textit{clase de alfabetiza\c{c}\~{a}}, followed by eight `` series''.
Primary education is mandatory for children aged 6-14. Secondary
education lasts three years. In principle, the age range of secondary
education students is 15-18. For each school, have geo-localized each
postal adress for a total of XX public primary and XX secondary schools
in the SPRM.

Our second source is information the 2010 Population Census of Brazil at
the \textit{clase de alfabetiza\c{c}\~{a}}, compiled and freely
distributed by the Brazilian Institute of Statistics (IBGE). This
spatial unit is equivalent to enumeration areas defined for surveying
purposes, and contain on average 400 households. There is a total of
30,815 areas in the SPMR. IBGE also provides the digital networks
containing the boundaries of the enumeration areas. From the census
microdata, we have data on the number of inhabitants by age for each
enumeration area. As the data was compiled in 2010 and we are interested
in the number of children in primary school age (ages 6-14), and
adolescents in secondary school age (ages 15-18), we calculate the count
of children in the age range 8-16 (who were 6 and 14 in 2008) and
adolescents in the age range 17-20 (aged 15-18 in 2008). The underlying
assumption is that the location of children and adolescents did not
change dramatically or systematically between 2008 and 2010, which is a
reasonable assumption given thatr residential mobility is relatively
low. using this procedure, we obtain a total number of children in
primary school age of 2,731,999 (approximately 13 percent of the
population) and a total of 1,730,252 tenneagers in secondary school age
(approximately 8 percent of the population).

Out last source is the 2007 Origin-Destination Household Travel Survey
(O-D Survey), carried out by the São Paulo Metropolitan Transport
Authority M\^{}\{e\}tro. we have information at the travel flow-level.
geographic coordinates are provided for the origin and destination of
each trip, as well as additional information such as trip purpose
(includng education), trip duration, number of transfers and mode. The
O-D Survey also provides the exact location of schools. The survey also
provides relevant information on individuals undertaking the trips, such
as their age, income range and educational status (including student).
We focus on trips made by primary and secondary schools with an
indicated motive ``education'' at the destination. The total number of
trips (excluding missing values) was 11,845. Of these, 5.5 percent were
multimodal. In those cases, we assing the mode of the trip leg with the
largest duration. We add all trip leg durations (in minutes) to obtain
the total trip duration.

@@

The cumulative-opportunity measure for mode \(M\) (Boisjoly G. 2016) is
defined as:

@@

\begin{itemize}
\tightlist
\item
  Descriptive statistics about modal choice and travel time by AEPs and
  level of income
\item
  Spatial Distribution of Schools and Transit-Dependent Students at the
  AEP Level
\end{itemize}

@@

\hyperdef{}{ref-allen1993accesibility}{\label{ref-allen1993accesibility}}
Allen, W Bruce, Dong Liu, and Scott Singer. 1993. ``Accesibility
Measures of US Metropolitan Areas.'' \emph{Transportation Research Part
B: Methodological} 27 (6). Elsevier: 439--49.

\hyperdef{}{ref-boisjoly2016}{\label{ref-boisjoly2016}}
Boisjoly G., El-Geneidy A. 2016. ``Is Complexity Better? Assessing
Accessibility Measures That Account for Daily Fluctuations in Transit
and Jobs Availability.'' 95th Annual Meeting of the Transportation
Research Board, Washington D.C., USA.

\hyperdef{}{ref-hernandez2015mobilities}{\label{ref-hernandez2015mobilities}}
Hernandez, Daniel Oviedo, and Helena Titheridge. 2015. ``Mobilities of
the Periphery: Informality, Access and Social Exclusion in the Urban
Fringe in Colombia.'' \emph{Journal of Transport Geography}. Elsevier.

\hyperdef{}{ref-ingram1971concept}{\label{ref-ingram1971concept}}
Ingram, David R. 1971. ``The Concept of Accessibility: A Search for an
Operational Form.'' \emph{Regional Studies} 5 (2). Taylor \& Francis:
101--7.

\hyperdef{}{ref-maia2016access}{\label{ref-maia2016access}}
Maia, Maria Leonor, Karen Lucas, Geraldo Marinho, Enilson Santos, and
Jessica Helena de Lima. 2016. ``Access to the Brazilian City---From the
Perspectives of Low-Income Residents in Recife.'' \emph{Journal of
Transport Geography}. Elsevier.

\hyperdef{}{ref-de2015changes}{\label{ref-de2015changes}}
Sá, Thiago Hérick de, Leandro Martin Totaro Garcia, Grégore Iven Mielke,
Fabiana Maluf Rabacow, and Leandro Fórnias Machado de Rezende. 2015.
``Changes in Travel to School Patterns Among Children and Adolescents in
the São Paulo Metropolitan Area, Brazil, 1997--2007.'' \emph{Journal of
Transport \& Health} 2 (2). Elsevier: 143--50.
